Bucket-Sorted Independent Sets for Algebraic Multigrid

### 2007-09-20 bsis_paper/paper.tex
git-svn-id: file:///home/rblake/svn/pamg/trunk@27 c7f3fe13-8c30-0410-9127-9a9c42fe009c

\begin{abstract}
\end{abstract}

@@

\subsection{Motivation}
\label{sec:motivation}
CLJP and its descendents select an independent set $D$ in each
iteration. Stated originally in Section~\ref{3:sec:CLJP}, the
condition for selecting $D$ is that all vertices $i \in D$ must
satisfy
\begin{equation}
\label{5:eqn:D-conditions}
w_i > w_j\;\textrm{for all } j \in \mathcal{N}_i.
\end{equation}
CLJP and CLJP-c rely on a search routine to locate locally maximal
weights and on a weight update routine to modify weights of vertices
connected to new $C$-points.

@@

In each iteration, the non-empty bucket with largest weight forms $D$
(see Corollary~\ref{5:cor:max-weight-indep-set}). These vertices are
assigned to the $C$-point set and removed from the data
structure. Vertex weight updates lead to corresponding updates to the
data structure, and new $F$-points are removed from the data
structure. These operations continue until all buckets are empty, at
which point the coarse-grid selection is
complete. Algorithms~\ref{5:alg:BSIS-setup},
\ref{5:alg:BSIS-indep-set-select}, and~\ref{5:alg:BSIS-weight-update}
outline the operations discussed above.
\begin{lstlisting}[caption={BSIS Data Structure Setup},label=5:alg:BSIS-setup]
%\textsc{BSIS-Setup}$(S)$%
for $i \in V$
  $bucketID \leftarrow (w_i-1)\cdot numColors + color_i$
  $bucket[bucketID].\textsc{Insert}(i)$
end
\end{lstlisting}
% \begin{algorithm}[!h]
% \caption{BSIS Data Structure Setup}
% \label{5:alg:BSIS-setup}
% \textsc{BSIS-Setup}$(S)$ \{
% \begin{algorithmic}[1]
% \FORALL{$i \in V$}
% \STATE $bucketID \leftarrow (w_i-1)\cdot numColors + color_i$
% \STATE $bucket[bucketID].\textsc{Insert}(i)$
% \ENDFOR
% \end{algorithmic}
% \}
% \end{algorithm}
\begin{lstlisting}[caption={Independent Set Selection},label=5:alg:BSIS-indep-set-select]
%\textsc{BSIS-Independent-Set-Selection}$(S)$%
return non-empty $bucket$ with largest $bucketID$
\end{lstlisting}
% \begin{algorithm}[!h]
% \caption{Independent Set Selection}
% \label{5:alg:BSIS-indep-set-select}
% \textsc{BSIS-Independent-Set-Selection}$(S)$ \{
% \begin{algorithmic}[1]
% \RETURN non-empty $bucket$ with largest $bucketID$
% \end{algorithmic}
% \}
% \end{algorithm}
\begin{lstlisting}[caption={BSIS Weight Update},label=5:alg:BSIS-weight-update]
%\textsc{BSIS-Weight-Update}$(S)$%
$bucketID \leftarrow (w_i-1)\cdot numColors + color_i$
$bucket[bucketID].\textsc{Remove}(i)$
$bucket[bucketID - numColors].\textsc{Insert}(i)$
\end{lstlisting}
%\begin{algorithm}[!h]
%\caption{BSIS Weight Update}
%\label{5:alg:BSIS-weight-update}
%\textsc{BSIS-Weight-Update}$(S)$ \{
%\begin{algorithmic}[1]
%\STATE $bucketID \leftarrow (w_i-1)\cdot numColors + color_i$
%\STATE $bucket[bucketID].\textsc{Remove}(i)$
%\STATE $bucket[bucketID - numColors].\textsc{Insert}(i)$
%\end{algorithmic}
%\}
%\end{algorithm}
\begin{figure}
  \begin{center}
    %\includegraphics[width=0.8\textwidth]{images/bucket-algorithm}
    \includegraphics[width=0.98\textwidth]{images/BSIS/initial}
    \caption{The BSIS data structure.}
    \label{5:fig:bucket-algorithm}
  \end{center}
\end{figure}

>>>>

\begin{abstract}
Independent set-based coarse-grid selection algorithms are defined by
policies regarding: weight initialization, independent set selection,
and weight update. We develop theory demonstrating that all algorithms
employing the same policies for each routine produce identical coarse
grids, regardless of the implementation. This result on coarse grid
invariance is applied to produce a new coarse-grid selection algorithm
that is substantially more efficient than an existing algorithm that
utilizes the same policies. Experimental results comparing two
versions of the new algorithm and an older algorithm are presented,
followed by a discussion on possible approaches to effectively
parallelize the new algorithms.
\end{abstract}

@@

\subsection{Motivation}
\label{sec:motivation}
CLJP and its descendents select an independent set $D$ in each
iteration. Recall, the condition for selecting $D$ is that all
vertices $i \in D$ must satisfy
\begin{equation}
\label{5:eqn:D-conditions}
w_i > w_j\;\textrm{for all } j \in \mathcal{N}_i.
\end{equation}
CLJP and CLJP-c rely on a search routine to locate locally maximal
weights and on a weight update routine to modify weights of vertices
connected to new $C$-points.

@@

In each iteration, the non-empty bucket with largest weight forms $D$
(see Corollary~\ref{cor:max-weight-indep-set}). These vertices are
assigned to the $C$-point set and removed from the data
structure. Vertex weight updates lead to corresponding updates to the
data structure, and new $F$-points are removed from the data
structure. These operations continue until all buckets are empty, at
which point the coarse-grid selection is
complete. Algorithms~\ref{5:alg:BSIS-setup},
\ref{5:alg:BSIS-indep-set-select}, and~\ref{5:alg:BSIS-weight-update}
outline the operations discussed above.
\begin{lstlisting}[caption={BSIS Data Structure Setup},label=5:alg:BSIS-setup]
%\textsc{BSIS-Setup}$(S)$%
for $i \in V$
  $bucketID \leftarrow (w_i-1)\cdot numColors + color_i$
  $bucket[bucketID].\textsc{Insert}(i)$
end
\end{lstlisting}
% \begin{algorithm}[!h]
% \caption{BSIS Data Structure Setup}
% \label{5:alg:BSIS-setup}
% \textsc{BSIS-Setup}$(S)$ \{
% \begin{algorithmic}[1]
% \FORALL{$i \in V$}
% \STATE $bucketID \leftarrow (w_i-1)\cdot numColors + color_i$
% \STATE $bucket[bucketID].\textsc{Insert}(i)$
% \ENDFOR
% \end{algorithmic}
% \}
% \end{algorithm}
\begin{lstlisting}[caption={Independent Set Selection},label=5:alg:BSIS-indep-set-select]
%\textsc{BSIS-Independent-Set-Selection}$(S)$%
return non-empty $bucket$ with largest $bucketID$
\end{lstlisting}
% \begin{algorithm}[!h]
% \caption{Independent Set Selection}
% \label{5:alg:BSIS-indep-set-select}
% \textsc{BSIS-Independent-Set-Selection}$(S)$ \{
% \begin{algorithmic}[1]
% \RETURN non-empty $bucket$ with largest $bucketID$
% \end{algorithmic}
% \}
% \end{algorithm}
\begin{lstlisting}[caption={BSIS Weight Update},label=5:alg:BSIS-weight-update]
%\textsc{BSIS-Weight-Update}$(S)$%
$bucketID \leftarrow (w_i-1)\cdot numColors + color_i$
$bucket[bucketID].\textsc{Remove}(i)$
$bucket[bucketID - numColors].\textsc{Insert}(i)$
\end{lstlisting}
%\begin{algorithm}[!h]
%\caption{BSIS Weight Update}
%\label{5:alg:BSIS-weight-update}
%\textsc{BSIS-Weight-Update}$(S)$ \{
%\begin{algorithmic}[1]
%\STATE $bucketID \leftarrow (w_i-1)\cdot numColors + color_i$
%\STATE $bucket[bucketID].\textsc{Remove}(i)$
%\STATE $bucket[bucketID - numColors].\textsc{Insert}(i)$
%\end{algorithmic}
%\}
%\end{algorithm}
\begin{figure}
  \begin{center}
    %\includegraphics[width=0.8\textwidth]{images/bucket-algorithm}
    \includegraphics[width=0.98\textwidth]{images/BSIS/initial}
    \caption{The BSIS data structure.}
    \label{5:fig:bucket-algorithm}
  \end{center}
\end{figure}
